Static and Dynamic Bethe-Salpeter Equations in the $T$-Matrix Approximation

### 2022-03-29 Response_Letter/Response_Letter.tex

%%% REVIEWER 1 %%%
\noindent \textbf{\large Authors' answer to Reviewer \#1}

{The manuscript reports neutral excitation energy calculations with the Bethe-Salpeter equation in the $T$-matrix approximation. 
The paper is well written and reports important new results.

However general conclusions like "BSE@GT performs best for few-electron systems whereas BSE@GW is better for larger systems" 
are drawn out of only 3 systems: H2, H2O and BeH2. None of the 3 can be considered as a "large" system.

I think the work would be much stronger with a more comprehensive benchmark set.}
\\
\alert{
We thanks the reviewer for these positive comments.
We are planning on performing a comprehensive benchmark in a forthcoming paper but we feel that it is currently outside the scope of the present paper which is dedicated to the introduction of a new theory and test calculations.
Accordingly to the reviewer's comment, we have soften our conclusion about the accuracy of BSE@GT for large systems while mentioning that a more comprehensive benchmark study would be required:
``The overall accuracy of the present scheme still needs to be assessed for larger systems (where the screening is known to be more important). 
For such purposes, a comprehensive benchmark study would be required and we are planning to do so in the future''
}

@@

\begin{enumerate}

\item 
{Eq. (3) shows the non-interacting Green's function. 
The word "non-interacting" is missing in the text (else $i$ and $a$ are not orbital indices). }
\\
\alert{
This is actually not the definition of the non-interacting Green's function, but the definition of the Green's function in the quasiparticle approximations [which can be defined at the Hartree-Fock, Kohn-Sham, or $GW$ level as mentioned just below Eq.~(3)].
Thank you for pointing out this misunderstanding; we have clarified this statement.
}

### 2022-03-29 Manuscript/Tmatrix.tex
update response letter and mansucript

\documentclass[aip,jcp,reprint,noshowkeys,superscriptaddress]{revtex4-1}
\usepackage{graphicx,dcolumn,bm,xcolor,microtype,multirow,amscd,amsmath,amssymb,amsfonts,physics,longtable,wrapfig,txfonts,siunitx}
\usepackage[version=4]{mhchem}

@@

%%%%%%%%%%%%%%%%%%%%%%
\section{Charged excitations}
\label{sec:charged}
%%%%%%%%%%%%%%%%%%%%%%
By definition, the one-body Green's function is \cite{Martin_2016}
\begin{equation}
\label{eq:G}
	G(\bx_1,\bx_2;\omega) 
	= \sum_i \frac{\SO{i}(\bx_1) \SO{i}(\bx_2)}{\omega - \e{i}{} - i\eta}	
	+ \sum_a \frac{\SO{a}(\bx_1) \SO{a}(\bx_2)}{\omega - \e{a}{} + i\eta}	
\end{equation}
where $\eta$ is a positive infinitesimal, and its nature is completely defined by the set of orbitals and corresponding energies that are used to build it.
For example, $G^{\HF}(\bx_1,\bx_2;\omega)$ is the Hartree-Fock (HF) Green's function built from the HF spinorbitals $\SO{p}^{\HF}(\bx)$ and energies $\e{p}{\HF}$.

@@

%%% FIGURE 1 %%%
\begin{figure*}[t]
	\includegraphics[width=\textwidth]{fig1}
	\caption{Singlet (left) and triplet (right) excitation energies (in \si{\eV}) of \ce{H2} as a function of the internuclear distance $R_{\ce{H-H}}$ (in \si{\angstrom}) computed at the FCI (black), BSE@$G_0W_0$ (blue), and BSE@$G_0T_0$ (red) levels with the cc-pVTZ basis.
	Raw data are reported in {\SupMat}.}
	\label{fig:H2}
\end{figure*}
%%% %%% %%% %%%

%%% FIGURE 2 %%%
\begin{figure}[t]
	\includegraphics[width=\linewidth]{fig2}
	\caption{HOMO and LUMO quasiparticle energies as well as HOMO-LUMO gap (in \si{\eV}) of \ce{H2} as a function of the internuclear distance $R_{\ce{H-H}}$ (in \si{\angstrom}) computed at the $G_0W_0$ (blue) and $G_0T_0$ (red) levels with the cc-pVTZ basis.
	Raw data are reported in {\SupMat}.}
	\label{fig:H2_gap}
\end{figure}
%%% %%% %%% %%%

%%% FIGURE 3 %%%
\begin{figure*}[t]
	\includegraphics[width=0.4\linewidth]{fig3a}
	\hspace{0.1\linewidth}
	\includegraphics[width=0.4\linewidth]{fig3b}
	\caption{
	Error with respect to FCI for the lowest singlet (left) and the lowest triplet (right) excitation energies of \ce{H2} as a function of the internuclear distance $R_{\ce{H-H}}$ (in \si{\angstrom}) computed within the static schemes (BSE@$G_0W_0$ and BSE@$G_0T_0$) and the dynamically-corrected schemes (dBSE@$G_0W_0$ and dBSE@$G_0T_0$) .
	The cc-pVTZ basis is employed for all calculations.
	Raw data are reported in {\SupMat}.}
	\label{fig:H2_dBSE}
\end{figure*}
%%% %%% %%% %%%

@@

%%% FIGURE 4 %%%
\begin{figure*}[t]
	\includegraphics[width=\textwidth]{fig4}
	\caption{Singlet (left) and triplet (right) excitation energies (in \si{\eV}) of \ce{BeH2} as a function of the distance $R_{\ce{Be-H}}$ (in \si{\angstrom}) computed at the FCI (black), BSE@$G_0W_0$ (blue), and BSE@$G_0T_0$ (red) levels with the cc-pVDZ basis.
	The dynamically-corrected BSE excitation energies are represented as thin lines for dBSE@$G_0W_0$ (blue) and dBSE@$G_0T_0$ (red).
	Raw data are reported in {\SupMat}.}
	\label{fig:BeH2}
\end{figure*}
%%% %%% %%% %%%

%%% FIGURE 5 %%%
\begin{figure}[t]
	\includegraphics[width=\linewidth]{fig5}
	\caption{HOMO and LUMO quasiparticle energies as well as HOMO-LUMO gap (in \si{\eV}) of \ce{BeH2} as a function of the internuclear distance $R_{\ce{Be-H}}$ (in \si{\angstrom}) computed at the BSE@$G_0W_0$ (blue) and BSE@$G_0T_0$ (red) levels with the cc-pVDZ basis.
	Raw data are reported in {\SupMat}.}
	\label{fig:BeH2_gap}
\end{figure}
%%% %%% %%% %%%

@@

%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:ccl}
%%%%%%%%%%%%%%%%%%%%%%
We have derived and implemented, for the first time, the static and dynamic Bethe-Salpeter equations when one considers $T$-matrix quasiparticle energies as well as a $T$-matrix-based kernel.
The performance of the static scheme and its perturbative dynamical correction have been assessed by computing the neutral excited states of several molecular systems.
Our results suggest that, in the context of the computation of molecular excitation energies, the BSE@$GT$ formalism performs best in few-electron systems where the electron density remains low.
\alert{The overall accuracy of the present scheme still needs to be assessed for larger systems (where the screening is known to be more important). 
For such purposes, a comprehensive benchmark study would be required and we are planning to do so in the future}
%However, it still outperforms conventional methods such CIS and TDHF for singlet states.

@@

\begin{equation}
	\label{eq:egval}
	\qty(\e{a}{}-\e{i}{}-\omega_S) X_{ia,S}+\sum_{jb}X_{jb,S}\mel{ib}{}{aj} + \sum_{jb}X_{jb,S}\widetilde{\cT}^c_{ib,aj}(\omega_S)=0
\end{equation}
where the spectral representation of the dynamical $T$-matrix is
\begin{equation}
	\widetilde{\cT}^c_{ib,aj}(\omega_S)
	= \frac{i}{2\pi}\int d\omega \lim_{\eta \to 0^+}\cT^c_{ib,aj}(\omega)e^{-i\omega\eta} \qty[\frac{1}{\omega_S-\omega+\e{j}{}+\e{i}{}+i\eta}+\frac{1}{\omega_S+\omega-\e{b}{}-\e{a}{}+i\eta}] 
 \label{eq:H2preso_eigen_2_T}
\end{equation}
with $\cT^c_{ib,aj}=\cT_{ib,aj}-\mel{ib}{}{aj}$ the correlation part of $\cT$. 
Equation \eqref{eq:egval} represents a non-linear eigenvalue equation to calculate the positive excitation energies of a system, which can be rewritten as
\begin{equation}
\label{eq:H2preso_eigen}
	\sum_{jb} A_{ia,jb}(\omega_S) X_{jb,S}=\omega_S X_{ia,S}
\end{equation}
with
\begin{equation}
	A_{ia,jb}(\omega_S)=(\e{a}{}-\e{i}{})\delta_{ij}\delta_{ab} + \mel{ib}{}{aj} + \widetilde{\cT}^c_{ib,aj}(\omega_S).
\end{equation}
If one drops the dynamical part $\widetilde{\cT}^c$, one ends up with the usual time-dependent Hartree-Fock (TDHF) equations. \cite{Dreuw_2005} To calculate the correlation contribution, one can employ Eq.~\eqref{eq:T} in Eq.~\eqref{eq:H2preso_eigen_2_T}, and, after integration over the frequency, one gets Eq.~\eqref{eq:dT}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\twocolumngrid
\section*{Data availability statement}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The data that supports the findings of this study are available within the article and its supplementary material.

>>>>

\documentclass[aip,jcp,reprint,noshowkeys,superscriptaddress]{revtex4-1}
\usepackage{graphicx,dcolumn,bm,xcolor,microtype,multirow,amscd,amsmath,amssymb,amsfonts,physics,longtable,wrapfig,txfonts,siunitx,float}
\usepackage[version=4]{mhchem}

@@

%%%%%%%%%%%%%%%%%%%%%%
\section{Charged excitations}
\label{sec:charged}
%%%%%%%%%%%%%%%%%%%%%%
By definition, \alert{in the quasiparticle approximation,} the one-body Green's function is \cite{Martin_2016}
\begin{equation}
\label{eq:G}
	G(\bx_1,\bx_2;\omega) 
	= \sum_i \frac{\SO{i}(\bx_1) \SO{i}(\bx_2)}{\omega - \e{i}{} - i\eta}	
	+ \sum_a \frac{\SO{a}(\bx_1) \SO{a}(\bx_2)}{\omega - \e{a}{} + i\eta}	
\end{equation}
where $\eta$ is a positive infinitesimal, and its nature is completely defined by the set of orbitals and corresponding energies that are used to build it.
For example, $G^{\HF}(\bx_1,\bx_2;\omega)$ is the Hartree-Fock (HF) Green's function built from the HF spinorbitals $\SO{p}^{\HF}(\bx)$ and energies $\e{p}{\HF}$.

@@

%%% FIGURE 1 %%%
\begin{figure*}
	\includegraphics[width=\textwidth]{fig1}
	\caption{Singlet (left) and triplet (right) excitation energies (in \si{\eV}) of \ce{H2} as a function of the internuclear distance $R_{\ce{H-H}}$ (in \si{\angstrom}) computed at the FCI (black), BSE@$G_0W_0$ (blue), and BSE@$G_0T_0$ (red) levels with the cc-pVTZ basis.
	Raw data are reported in {\SupMat}.}
	\label{fig:H2}
\end{figure*}
%%% %%% %%% %%%

%%% FIGURE 2 %%%
\begin{figure}
	\includegraphics[width=\linewidth]{fig2}
	\caption{HOMO and LUMO quasiparticle energies as well as HOMO-LUMO gap (in \si{\eV}) of \ce{H2} as a function of the internuclear distance $R_{\ce{H-H}}$ (in \si{\angstrom}) computed at the $G_0W_0$ (blue) and $G_0T_0$ (red) levels with the cc-pVTZ basis.
	Raw data are reported in {\SupMat}.}
	\label{fig:H2_gap}
\end{figure}
%%% %%% %%% %%%

%%% FIGURE 3 %%%
\begin{figure*}
	\includegraphics[width=0.4\linewidth]{fig3a}
	\hspace{0.1\linewidth}
	\includegraphics[width=0.4\linewidth]{fig3b}
	\caption{
	Error with respect to FCI for the lowest singlet (left) and the lowest triplet (right) excitation energies of \ce{H2} as a function of the internuclear distance $R_{\ce{H-H}}$ (in \si{\angstrom}) computed within the static schemes (BSE@$G_0W_0$ and BSE@$G_0T_0$) and the dynamically-corrected schemes (dBSE@$G_0W_0$ and dBSE@$G_0T_0$) .
	The cc-pVTZ basis is employed for all calculations.
	Raw data are reported in {\SupMat}.}
	\label{fig:H2_dBSE}
\end{figure*}
%%% %%% %%% %%%

@@

%%% FIGURE 4 %%%
\begin{figure*}
	\includegraphics[width=\textwidth]{fig4}
	\caption{Singlet (left) and triplet (right) excitation energies (in \si{\eV}) of \ce{BeH2} as a function of the distance $R_{\ce{Be-H}}$ (in \si{\angstrom}) computed at the FCI (black), BSE@$G_0W_0$ (blue), and BSE@$G_0T_0$ (red) levels with the cc-pVDZ basis.
	The dynamically-corrected BSE excitation energies are represented as thin lines for dBSE@$G_0W_0$ (blue) and dBSE@$G_0T_0$ (red).
	Raw data are reported in {\SupMat}.}
	\label{fig:BeH2}
\end{figure*}
%%% %%% %%% %%%

%%% FIGURE 5 %%%
\begin{figure}
	\includegraphics[width=\linewidth]{fig5}
	\caption{HOMO and LUMO quasiparticle energies as well as HOMO-LUMO gap (in \si{\eV}) of \ce{BeH2} as a function of the internuclear distance $R_{\ce{Be-H}}$ (in \si{\angstrom}) computed at the BSE@$G_0W_0$ (blue) and BSE@$G_0T_0$ (red) levels with the cc-pVDZ basis.
	Raw data are reported in {\SupMat}.}
	\label{fig:BeH2_gap}
\end{figure}
%%% %%% %%% %%%

@@

%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:ccl}
%%%%%%%%%%%%%%%%%%%%%%
We have derived and implemented, for the first time, the static and dynamic Bethe-Salpeter equations when one considers $T$-matrix quasiparticle energies as well as a $T$-matrix-based kernel.
The performance of the static scheme and its perturbative dynamical correction have been assessed by computing the neutral excited states of several molecular systems.
Our results suggest that, in the context of the computation of molecular excitation energies, the BSE@$GT$ formalism performs best in few-electron systems where the electron density remains low.
\alert{The overall accuracy of the present scheme still needs to be assessed for larger systems (where the screening is known to be more important). 
For such purposes, a comprehensive benchmark study would be required and we are planning to do so in the future.}
%However, it still outperforms conventional methods such CIS and TDHF for singlet states.

@@

\begin{equation}
	\label{eq:egval}
	\qty(\e{a}{}-\e{i}{}-\omega_S) X_{ia,S}+\sum_{jb}X_{jb,S}\mel{ib}{}{aj} + \sum_{jb}X_{jb,S}\widetilde{\cT}^c_{ib,aj}(\omega_S)=0
\end{equation}
where the spectral representation of the dynamical $T$-matrix is
\begin{equation}
	\widetilde{\cT}^c_{ib,aj}(\omega_S)
	= \frac{i}{2\pi}\int d\omega \lim_{\eta \to 0^+}\cT^c_{ib,aj}(\omega)e^{-i\omega\eta} \qty[\frac{1}{\omega_S-\omega+\e{j}{}+\e{i}{}+i\eta}+\frac{1}{\omega_S+\omega-\e{b}{}-\e{a}{}+i\eta}] 
 \label{eq:H2preso_eigen_2_T}
\end{equation}
with $\cT^c_{ib,aj}=\cT_{ib,aj}-\mel{ib}{}{aj}$ the correlation part of $\cT$. 
Equation \eqref{eq:egval} represents a non-linear eigenvalue equation to calculate the positive excitation energies of a system, which can be rewritten as
\begin{equation}
\label{eq:H2preso_eigen}
	\sum_{jb} A_{ia,jb}(\omega_S) X_{jb,S}=\omega_S X_{ia,S}
\end{equation}
with
\begin{equation}
	A_{ia,jb}(\omega_S)=(\e{a}{}-\e{i}{})\delta_{ij}\delta_{ab} + \mel{ib}{}{aj} + \widetilde{\cT}^c_{ib,aj}(\omega_S).
\end{equation}
If one drops the dynamical part $\widetilde{\cT}^c$, one ends up with the usual time-dependent Hartree-Fock (TDHF) equations. \cite{Dreuw_2005} To calculate the correlation contribution, one can employ Eq.~\eqref{eq:T} in Eq.~\eqref{eq:H2preso_eigen_2_T}, and, after integration over the frequency, one gets Eq.~\eqref{eq:dT}.
\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\twocolumngrid
\section*{Data availability statement}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The data that supports the findings of this study are available within the article and its supplementary material.
